\bf Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-07-24 manuscript/main.tex
residuals are centered

	\subsection{Step 2: Computing Residualized Predictors via Hybrid-on-Scalar Regression} \label{subsec: Step 2: Computing Predictors via Hybrid-on-Scalar Regression}
The $l$-th iteration's second step involves residualization of both predictors and responses. We first compute the individual PLS score:  
\begin{equation}\label{def:plsscore}
	\hat{\rho}^{[l]}_i := \langle \widetilde{W}_i^{[l]}, \, \hat{\xi}^{[l]} \rangle_{\mathcal{H}},
\end{equation}
utilizing the estimated PLS component direction $\widehat{\xi}^{[l]}$ obtained from Propositions \ref{proposition:eigen_regul} and \ref{proposition:linear_regul}. 
Using these PLS scores as scalar predictor observations, the $(l+1)$-th iteration's residualized response is computed as:
\begin{equation}\label{algorithm_step:residualization_response}
	Y_i^{[l+1]} := Y_i^{[l]} - 
	\frac{
		( \mathbf{y}^{[l]} )^\top  \hat{\boldsymbol{\rho}}^{[l]}
	}{
		( \hat{\boldsymbol{\rho}}^{[l]} )^\top  \hat{\boldsymbol{\rho}}^{[l]}
	}
	\widehat{\rho}_{il},
\end{equation}
which is equivalent to using the least squares estimate from regressing $Y_i^{[l]}$ on $\widehat{\rho}_{il}$,
a procedure identical to that found in the naive PLS algorithm (Section \ref{sec: Naive PLS Algorithm}).

@@

 \begin{lemma}[Closed-form  solution]\label{proposition:closed_form_orthgonalization}
 	Let us denote
 	\begin{equation*}
 		 		\hat{\boldsymbol{\rho}}^{[l]} := (\hat{\rho}^{[l]}_1, \ldots, \hat{\rho}^{[l]}_n)^\top
 	\quad\text{and} \quad
 			\bar{\rho}^{[l]} := (\hat{\boldsymbol{\rho}}^{[l]})^\top \hat{\boldsymbol{\rho}}^{[l]}.
 	\end{equation*}
 	The solution to the optimization problem \eqref{def:argmin_pensse}, denoted
 	$\hat{\delta}^{[l]}   \in \tilde{\mathcal{H}}$, has basis coefficients and scalar parts given by:
\begin{equation*}
	\hat{\boldsymbol{\pi}}_j^{[l]}
	=
	\bigl(
		\bar{\rho}^{[l]}
  B
	+
	\tau
	B^{\prime \prime}
	\bigr)^{-1}
	  B
	(\Theta_j^{[l]})^\top,
	~
	j = 1, \ldots, K,
	%%%%%%%%%%%%%%%%%%%%%%%%
	\quad
	\hat{\boldsymbol{\chi} }
	=
	(\mathbf{Z}^{[l]})^\top  \hat{\boldsymbol{\rho}}.
\end{equation*}
 \end{lemma}
 Proof of Proposition \ref{proposition:closed_form_orthgonalization} is provided in Appendix \ref{section:proof:proposition:closed_form_orthgonalization}.

 The $(l+1)$-th iteration's residualized predictor is then computed as:
\begin{equation}\label{algorithm_step:residualization_predictor}
	\widetilde{W}_i^{[l+1]} 
	 = 
	 \widetilde{W}_i^{[l]}  - \widehat{\rho}_{il}  \widehat{ {\delta}}_{l}.
	%%%%%%%%%%%
\end{equation}

@@

\section{Proof of Proposition \ref{proposition:closed_form_orthgonalization} }\label{section:proof:proposition:closed_form_orthgonalization}
For notational simplicity, we omit the iteration superscript $[l]$.
	Using the basis expansion coefficient notation from \eqref{def:truncated_zeta_and_delta}, \eqref{def:coef_vector_form}, \eqref{def:theta}, and \eqref{def:B_matrix}, the full vector of PLS scores $\hat{\boldsymbol{\rho}}  = (\hat{\rho}_1, \ldots, \hat{\rho}_n)^\top$ 
	where
	$\hat{\rho}_1$ is defined in \eqref{def:plsscore},
	is computed through the following matrix multiplication:
	\begin{equation*}
		 \hat{\boldsymbol{\rho}}^\top = \sum_{k=1}^K ( \hat{\boldsymbol{\gamma}}_j )^\top B \, \Theta_j^\top + \boldsymbol{\zeta}^\top \mathbf{Z}^\top.
	\end{equation*}
The criterion $\operatorname{PENSSE}( \delta)$ can be decomposed as follows:

@@

This completes the proof of Proposition \ref{proposition:closed_form_orthgonalization}.

>>>>

	\subsection{Step 2: Computing Residualized Predictors via Hybrid-on-Scalar Regression} \label{subsec: Step 2: Computing Predictors via Hybrid-on-Scalar Regression}
The $l$-th iteration's second step involves residualization of both predictors and responses. We first compute the individual PLS score:  
\begin{equation}\label{def:plsscore}
	\hat{\rho}^{[l]}_i := \langle \widetilde{W}_i^{[l]}, \, \hat{\xi}^{[l]} \rangle_{\mathcal{H}},
\end{equation}
using the estimated PLS component direction $\widehat{\xi}^{[l]}$ obtained from Propositions \ref{proposition:eigen_regul} and \ref{proposition:linear_regul}. 
Since   $\widetilde{W}_i^{[l]}$ are assumed to have a sample mean of zero, these PLS scores will also have a sample mean of zero.
Using these PLS scores as scalar predictor observations, the $(l+1)$-th iteration's residualized response is computed as:
\begin{equation}\label{algorithm_step:residualization_response}
	Y_i^{[l+1]} := Y_i^{[l]} - 
	\frac{
		( \mathbf{y}^{[l]} )^\top  \hat{\boldsymbol{\rho}}^{[l]}
	}{
		( \hat{\boldsymbol{\rho}}^{[l]} )^\top  \hat{\boldsymbol{\rho}}^{[l]}
	}
	\widehat{\rho}_{il},
\end{equation}
which is equivalent to using the least squares estimate from regressing $Y_i^{[l]}$ on $\widehat{\rho}_{il}$,
a procedure identical to that found in the naive PLS algorithm (Section \ref{sec: Naive PLS Algorithm}). 
Since $Y_i^{[l]}$ are assumed to have a sample mean of zero, these residuals $Y_i^{[l+1]}$   also have a sample mean of zero.

@@

 \begin{lemma}[Closed-form  solution]\label{proposition:closed_form_orthgonalization}
 	Let us denote
 	\begin{equation*}
 		 		\hat{\boldsymbol{\rho}}^{[l]} := (\hat{\rho}^{[l]}_1, \ldots, \hat{\rho}^{[l]}_n)^\top
 	\quad\text{and} \quad
 			\bar{\rho}^{[l]} := (\hat{\boldsymbol{\rho}}^{[l]})^\top \hat{\boldsymbol{\rho}}^{[l]}.
 	\end{equation*}
 	The solution to the optimization problem \eqref{def:argmin_pensse}, denoted
 	$\hat{\delta}^{[l]}   \in \tilde{\mathcal{H}}$, has basis coefficients and scalar parts given by:
\begin{equation*}
	\hat{\boldsymbol{\pi}}_j^{[l]}
	=
	\bigl(
		\bar{\rho}^{[l]}
  B
	+
	\tau
	B^{\prime \prime}
	\bigr)^{-1}
	  B
	(\Theta_j^{[l]})^\top
	\hat{\boldsymbol{\rho}}
	,
	~
	j = 1, \ldots, K,
	%%%%%%%%%%%%%%%%%%%%%%%%
	\quad
	\hat{\boldsymbol{\chi} }
	=
	(\mathbf{Z}^{[l]})^\top  \hat{\boldsymbol{\rho}}.
\end{equation*}
 \end{lemma}
 Proof of Lemma \ref{proposition:closed_form_orthgonalization} is provided in Appendix \ref{section:proof:proposition:closed_form_orthgonalization}.

 The $(l+1)$-th iteration's residualized predictor is then computed as:
\begin{equation}\label{algorithm_step:residualization_predictor}
	\widetilde{W}_i^{[l+1]} 
	 = 
	 \widetilde{W}_i^{[l]}  - \widehat{\rho}_{il}  \widehat{ {\delta}}_{l}.
	%%%%%%%%%%%
\end{equation}
 Since $\widetilde{W}_i^{[l]} $ are assumed to have a sample mean of zero, these residuals $\widetilde{W}_i^{[l+1]}$   also have a sample mean of zero.

@@

\section{Proof of Lemma \ref{proposition:closed_form_orthgonalization} }\label{section:proof:proposition:closed_form_orthgonalization}
For notational simplicity, we omit the iteration superscript $[l]$.
	Using the basis expansion coefficient notation from \eqref{def:truncated_zeta_and_delta}, \eqref{def:coef_vector_form}, \eqref{def:theta}, and \eqref{def:B_matrix}, the full vector of PLS scores $\hat{\boldsymbol{\rho}}  = (\hat{\rho}_1, \ldots, \hat{\rho}_n)^\top$ 
	where
	$\hat{\rho}_1$ is defined in \eqref{def:plsscore},
	is computed through the following matrix multiplication:
	\begin{equation*}
		 \hat{\boldsymbol{\rho}}^\top = \sum_{k=1}^K ( \hat{\boldsymbol{\gamma}}_j )^\top B \, \Theta_j^\top + \boldsymbol{\zeta}^\top \mathbf{Z}^\top.
	\end{equation*}
The criterion $\operatorname{PENSSE}( \delta)$ can be decomposed as follows:

@@

This completes the proof of Lemma \ref{proposition:closed_form_orthgonalization}.
